### 2016-08-25 section/3_redundancy.tex
shrink through section 3 and 4

\subsubsection{Design overview}
We will compare the computation of different iterations
from one loop instance and that of different loop instances from 
one static loop,
and judge whether there is redundancy.
Specifically, several questions need to be answered.

\emph{What to compare?}
Given two loop iterations (or loop instances) $c_1$ and $c_2$, since they 
originate from the same source code $C$, 
a naive approach is to record the value read by
every memory read instruction in $c_1$ and $c_2$. 
We will know whether $c_1$ and $c_2$ are redundant with each other by comparing
these values.
Of course, we could do better by checking fewer instructions, as
some of these values are determined by values read earlier in $c_1$ and/or 
$c_2$. Informally speaking, we only need to compare \textit{input}
values of $c_1$ and $c_2$ to decide whether they are redundant with each other. 
We will present our formal definition of \textit{input} values, and
how we identify and record them in
Section \ref{sec:dependence}.

@@

Specifically, \Tool first identifies side-effect instructions in $C$,
similar with that in Section \ref{sec:s_workless}. 
It then conducts 
backward static slicing 
from these instructions, considering both control and data
dependency. For every memory read instruction $r$ 
(\texttt{read(v)}) that static slicing encounters, \Tool checks whether $r$
satisfies either one of the following two conditions. 
If $r$ does, it is marked as
an \textit{input} instruction; otherwise, \Tool continues growing the
slice beyond $r$.
The two conditions are:
(1) the value of $v$ is defined outside $C$;
(2) $v$ is a heap or global variable.
The rationale for the first condition is that slicing outside $C$ is
unnecessary for redundancy judgement among instances of $C$.
The rationale for the second condition is that tracking data-dependency
through heap or global variables is complicated in multi-threaded C/C++
programs.

The analysis for cross-iteration and cross-loop redundancy analysis is similar
--- simply replacing $C$ in the above algorithm with one iteration or
the whole loop body.

@@

Our analysis considers function calls inside $C$ --- slicing is conducted
for return values of callees and side-effect instructions
inside callees. 
%In order to reduce the number of memory read we need to record inside callees, 
%we calculate previous writes conducted to the same address for each local-structure read and local-array read. 
%Since local structures and local arrays are mainly used in the declared functions, 
%it is easy to know whether read or write conducted on them are applied to the same address. 
We omit encountered constant values through slicing, because constant values will not affect the redundancy judgement.

Note that another possible approach is to simply record and compare what
are written by the side-effect instructions in $C$.
We did not choose this approach because we believe it is not sufficient
to judge redundancy. Consider a loop that goes through a character array 
to check if there exists any special characters in the array
and returns a boolean
variable denoting its checking result. Running this loop on 
different arrays is not redundant, but could easily be judged as redundant
when only comparing the loop side-effect (e.g., always returning 
\texttt{FALSE}).

@@

%TODO why not directly judge side-effect values

\paragraph{Example}
Figure \ref{fig:Apache34464} shows a simplified code-snippet from Apache,
containing a cross-loop redundancy bug. 
The \texttt{for} loop starting on line 7 
searches a string \texttt{source} for a target sub-string 
\texttt{target}. Since the outer loop, which starts on line 2, appends one 
character to \texttt{source} in every iteration (line 30), 
every execution of the \texttt{for} loop is working on a similar 
\texttt{source} string from its previous execution, with a lot of redundancy.

\begin{figure}
  \centering
  \lstset{basicstyle=\ttfamily\fontsize{8}{8}\selectfont,
     morekeywords={+},keepspaces=true,numbers=left}
  \mbox{\lstinputlisting[mathescape,boxpos=t]{figures/Apache34464.c}}
  \caption{A cross-loop redundant bug in Apache}
  \label{fig:Apache34464}
\end{figure}

Take the simple loop at line 10 of Figure~\ref{fig:Apache34464} as an example.
Its only side-effect is to update \texttt{i}.
For cross-loop redundancy analysis, we will consider the whole loop and 
identify these as \textit{input}s:
every \texttt{source[i]} read inside the loop;
the initial value of \texttt{i} defined outside the loop;
\texttt{max} defined outside the loop; 
and \texttt{first} defined outside the loop.
For cross-iteration redundancy analysis, the \textit{input}s for each
iteration is slightly different:
value \texttt{i} defined in previous iteration, 
the \texttt{source[i]} read in that iteration, 
and two values defined outside the loop, \texttt{max} and \texttt{first}.

Take the big loop at line 7 of Figure~\ref{fig:Apache34464} as an example.
The \textit{input}s for the whole loop include:
three values defined outside the loop
(\texttt{max}, \texttt{first}, \texttt{targetLen}) and every
\texttt{source[.]} and \texttt{target[.]} read inside the loop.
This set is already smaller than the set of all data read by the loop.
We will further shrink this set, removing invariant and others, 
in Section \ref{sec:perf}.

@@

After identifying \textit{inputs} using static analysis, 
\Tool then instruments the program to record \textit{inputs} values at run time.
%Specifically, we will assign
%a unique ID for each source instruction, and a pair of $<InstID, Value>$
%will be recorded at run time with the execution of a source instruction.
The run-time trace will also include delimiters and meta information that allows
trace analysis to differentiate values recorded from different instructions, 
loop
iterations, loop instances, and so on.
%TODO
%All values defined outside the loop is recorded in block O. 
%We record a delimiter, loop instance number and values defined in previous iterations in block H1'.
%Since we only consider redundant iterations from the same loop instance,  
%we do not record values defined outside the loop for cross-iteration redundancy. 
Once the trace under problematic workload is collected either during
off-line debugging or production runs, 
\Tool will process the trace and decide
whether the loops under study contain cross-iteration or 
cross-loop redundancy. We will first present our high-level algorithms, followed
by the exact implementation in our prototype.

\paragraph{High-level algorithms}
For cross-iteration redundancy, we need to answer two questions. 
First, how to judge whether
two iterations are doing redundant work?
Second, is a (dynamic) loop problematic when it contains only few 
iterations that are
redundant with each other?

Our answer to these two questions stick to one principle: there should be
sufficient amount of redundant computation to make a loop likely culprit
for a user-perceived performance problem and to make itself worthwhile to
get optimized by developers. Consequently,
for the first question, \Tool takes a strict 
definition --- only iterations that are doing exactly the same computation are
considered redundant. Since one iteration may not contain too much computation,
a weaker definition here may lead to many false positives. For the 
second question, we believe there should be a threshold. In our current 
prototype,
when the number of distinct iterations is less than half of the total iterations, 
we consider the loop to be cross-iteration redundant. 
%when the number of distinct iterations is less than half of the total iterations 
%iterations/(distinct iterations) = 2

@@

For example, Figure~\ref{fig:Apache37184} shows a loop from Apache-Ant that
contains cross-iteration redundancy.
Under the problem triggering input, there are
only few distinct listeners, and most of the
loop iterations in \texttt{fireMessageLoggedEvent}
are doing redundant work. 
%The problem for this bug is that developers say that the performance 
%loss happens in fireMessageLoggedEvent, but in the code I found, function
%messageLogged does not contain anything, and it is an empty function.
%I guess either developers do not have correct understanding of this bug,
%or I did not find the correct codes. 
}

For cross-loop redundancy, we need to answer similar questions:
how to judge whether two loop instances are doing redundant work?
%should they contain exactly the same number of iterations and doing exactly
%the same computation in each iteration?
%Second, is a (static) loop problematic when it contains two instances that
%are redundant with each other?

Following the same principle discussed above, we have slightly different
answers here.
We do not require two loop instances to
conduct exactly the same computation to be considered redundant. The rationale
is that a whole loop instance contains a lot of computation, much more than
one iteration in general. Even if only part of its computation
is redundant, it could still be the root-cause of a user-perceived performance
problem and worth developers' attention.

In practice, we rarely see cases where different loop 
instances are doing exactly the same computation.
In cross-loop redundancy examples shown in 
Figure~\ref{fig:Mozilla477564} and
Figure~\ref{fig:Apache34464}, each loop instance is doing similar, but not
exactly the same, work from its previous instance.

@@

The key of this implementation is
to judge whether two dynamic loop instances $l_1$ and $l_2$ 
are redundant or not.
The challenge is that $l_1$ and $l_2$ may have executed different numbers of
iterations; in different iterations, different sets of input instructions
may have executed. Therefore, we cannot simply chain values from different
input instructions and iterations together and compare two data sequences.
Instead, we decide to check the redundancy for each input instruction across
$l_1$ and $l_2$ first, and then use the average \textit{redundancy rate} of
all input instructions as the \textit{cross-loop redundancy rate} between
$l_1$ and $l_2$.

We calculate the redundancy for one input instruction $I$ by normalizing the
edit-distance between the two sequences of values returned by $I$ in the two
loop instances. The exact formula is the following:

@@

\[
\Scale[0.8]{
	Redundancy(I) =  \frac{dist(SeqA,SeqB) - (len(SeqA) - len(SeqB))}{len(SeqA)}
}
\]

Here, $SeqA$ and $SeqB$ represent the two value sequences corresponding to $I$
from two loop instances, with $SeqA$ being the longer sequence.
$dist$ means edit distance, and $len$ means the length of a value sequence.
Since the edit distance is at least the length-difference between the
two sequences and at most the length of the longer sequence, we use the 
subtraction and division shown in the formula above to normalize the
redundancy value.

@@

\paragraph{Cross-iteration redundancy analysis}
Our high-level sampling strategy is straightforward:
randomly decide at the
beginning of every iteration whether to track the values returned by
input instructions in this iteration.

The implementation is similar with previous sampling work 
\cite{liblit03,liblit05}.
Specifically, we create a clone of the original
loop iteration code, including functions called by the loop directly or
indirectly, and insert value-recording instructions along the
cloned copy. 
We then insert a code snippet that conducts random decision to
the beginning of a loop iteration. 
Two variables \texttt{CurrentID}, which is initialized as 0, 
and \texttt{NextSampleID}, which is initialized by a random integer, 
are maintained
in this code snippet. \texttt{CurrentID} is increased by 1
for each iteration. When it matches \texttt{NextSampleID}, the control
flow jumps to the value-recording clone of the loop iteration and the 
\texttt{NextSampleID} is increased by a random value. Different sampling
sparsity setting will determine the range from which the random value is
generated.

@@

\paragraph{Cross-loop redundancy analysis} 
We randomly decide at the beginning
of every loop instance whether to track values for this instance. 
Since we will need to compare two consecutive loop
instances for redundancy, once we decide to sample one loop instance, we will
make sure to sample the immediate next loop instance too.
The implementation is straightforward by cloning the whole loop and making
sampling decisions in the loop pre-headers.
%; (2)
%sampling is conducted when \texttt{CurrentID} equals either
%\texttt{NextSampleID} or \texttt{NextSampleID}$+1$, with \texttt{NextSampleID}
%increased by a random value in the latter case.

@@

First, we avoid tracking multiple read instructions that we can statically 
prove to return the same value.
Since we implement \Tool in LLVM, we leverage the SSA construction and the
\textit{mem2reg} pass in LLVM to avoid unnecessary tracking of
stack variables.
For example, the read of \texttt{max} on line 10 of 
Figure \ref{fig:Apache34464} is an \textit{input} instruction of the 
\texttt{while} loop on the same line. LLVM identifies it as a loop invariant
and lifts the read of \texttt{max}
out of the loop. Consequently, \Tool only records
its value once during each loop, instead of each iteration.
LLVM is conservative in lifting heap/global variables out of
a loop, for fear of changing the location of potential exceptions, etc.
\Tool conducts a best-effort
loop invariant analysis for heap/global variables.
For example, \Tool identifies that the values of
\texttt{aNode->localName} and \texttt{aNode->namespaceURI} do not change
throughout one loop instance
in Figure~\ref{fig:Mozilla477564}, and hence only traces them once
outside the loop. This analysis is sound but not complete.
\Tool may miss some loop invariants. For in-house debugging, since we know
the problem-triggering inputs do not lead to exceptions, this optimization
is safe. For production-run usage, users can turn off this heap/global
variable related optimization.

Second, we leverage the scalar evolution analysis in LLVM to remove the 
monitoring to some loop-induction related variables.
The scalar evolution analysis can tell which variables are
loop-induction variables (e.g., \texttt{i} on line 10
of Figure~\ref{fig:Apache34464}) and what are their strides. 
We then use this information in two ways. 
One is for cross-iteration redundancy checking.
If a loop iteration's input set contains a loop-induction
variable, we know different iterations' inputs would be different and 
hence conclude that there is no cross-iteration redundancy without any
run-time analysis. 
The other is for cross-loop redundancy checking.
When the address of a memory read is a loop induction variable, such as
\texttt{source[i]} on line 10 of Figure~\ref{fig:Apache34464},
we only record the address range at run time (i.e., the start and the
end), instead of every value returned by the memory read.
This optimization could lead to false positives ---
different loop instances may work on variables read from
similar memory locations but with different values.
We did not encounter such false positives in our experiments.

\subsection{Fix Strategy Recommendation}
\label{sec:redundant_fix}
\Tool suggests how to fix an inefficient loop to developers.
This process is straightforward following our taxonomy discussion in 
Section \ref{sec:study}, as there is often only one or two natural fix 
strategies for each sub-category of root causes.

Once a loop is identified as \emph{resultless}, \Tool will suggest
the loop to be completely deleted, in case of 0$^*$ resultless, or
conditionally skipped, in case of [0$|$1]*; \Tool will suggest a 
data-structure change in case of 0*1? loops.

When a loop is identified as \emph{redundant}, \Tool conducts
extra analysis to
decide whether batching or memoization should be suggested to fix the
redundancy problem.

For cross-iteration redundancy, \Tool suggests batching if the
redundancy is related to I/O operations and suggests memoization otherwise.
Specifically, 
when the only side effect of a loop is from I/O operations
and the same statement(s) is executed in every loop iteration, \Tool reports 
this
as an I/O related redundancy problem and suggests batching as a potential fix
strategy. Otherwise, \Tool suggests memoization.

For cross-loop redundancy, whether to use memoization or batching often
depends on which strategy is cheaper.
\Tool uses a simple
heuristic. If 
the side effect of each loop instance is to update 
a constant number of memory locations, like the 
buggy loop in Figure~\ref{fig:Mozilla477564} and Figure~\ref{fig:Apache34464}, 
we recommend memoization. Instead, if the side effect is to update
a sequence of memory locations, with the number of locations increasing
with the workload, memoization is unlikely to save much and
hence batching is suggested.

In addition to the above high-level fix strategies, \Tool can also provide some 
detailed information to help developers design the patch.
For example, in case of memoization fix strategy,
since \Tool knows what exactly are the memory reads that return the same values
again and again, it can help decide what to memoize; in case of
I/O-related batching, \Tool points out the exact I/O operations that should
be considered batching; and so on. 
Generating complete patches is out of the scope of \Tool.

%TODO what does batching mean for cross-loop redundancy???

>>>>

We will compare the computation of different iterations
from one loop instance and that of different loop instances from 
one static loop,
and judge whether there is redundancy.
Specifically, several questions need to be answered.

\emph{What to compare?}
Given two loop iterations (or loop instances) $c_1$ and $c_2$, since they 
originate from the same source code $C$, 
a naive approach is to record and compare the value read by
every memory read instruction in $c_1$ and $c_2$. 
%We will know whether $c_1$ and $c_2$ are redundant with each other by comparing
%these values.
We could do better by checking fewer instructions, as
some of these values are determined by values read earlier in $c_1$ and
$c_2$. Informally speaking, we only need to compare \textit{input}
values of $c_1$ and $c_2$ to decide whether they are redundant with each other. 
We will present formal definition of \textit{input} and the detailed algorithms
in Section \ref{sec:dependence}.

@@

Specifically, \Tool first identifies side-effect instructions in $C$,
similar with that in Section \ref{sec:s_workless}. 
It then conducts 
backward static slicing 
from these instructions, considering control and data
dependency. For every memory read $r$ 
(\texttt{read(v)}) that static slicing encounters, \Tool checks whether $r$
satisfies either one of the following two conditions. 
If $r$ does, it is marked as
an \textit{input} instruction; otherwise, \Tool continues growing the
slice beyond $r$.
The two conditions are:
(1) the value of $v$ is defined outside $C$;
(2) $v$ is a heap or global variable.
The rationale for the first condition is that slicing outside $C$ is
unnecessary for redundancy judgement among instances of $C$.
The rationale for the second condition is that tracking data-dependency
through heap or global variables is complicated in multi-threaded C/C++
programs.
The analysis for cross-iteration and cross-loop redundancy analysis is similar
--- simply replacing $C$ in the above algorithm with one iteration or
the whole loop body.

@@

Our analysis considers function calls inside $C$ --- slicing is conducted
for return values of callees and side-effect instructions
inside callees. 
%In order to reduce the number of memory read we need to record inside callees, 
%we calculate previous writes conducted to the same address for each local-structure read and local-array read. 
%Since local structures and local arrays are mainly used in the declared functions, 
%it is easy to know whether read or write conducted on them are applied to the same address. 
We omit encountered constant values through slicing, as they do not affect redundancy judgement.

\comment{
Note that another possible approach is to simply record and compare what
are written by the side-effect instructions in $C$.
We did not choose this approach because we believe it is not sufficient
to judge redundancy. Consider a loop that goes through a character array 
to check if there exists any special characters in the array
and returns a boolean
variable denoting its checking result. Running this loop on 
different arrays is not redundant, but could easily be judged as redundant
when only comparing the loop side-effect (e.g., always returning 
\texttt{FALSE}).
}

@@

\begin{figure}
  \centering
  \lstset{basicstyle=\ttfamily\fontsize{8}{8}\selectfont,
     morekeywords={+},keepspaces=true,numbers=left}
  \mbox{\lstinputlisting[mathescape,boxpos=t]{figures/Apache34464.c}}
  \caption{A cross-loop redundant bug in Apache (The \texttt{for} loop on line 7 
searches a string \texttt{source} for a target sub-string 
\texttt{target}. Since the outer loop on line 2 appends one 
character to \texttt{source} in every iteration (line 30), 
the \texttt{for} loop is always working on a similar 
\texttt{source} from its previous execution, with a lot of redundancy.) 
}
  \label{fig:Apache34464}
\end{figure}

\comment{
Take the simple loop at line 10 of Figure~\ref{fig:Apache34464} as an example.
Its only side-effect is to update \texttt{i}.
For cross-loop redundancy analysis, we will consider the whole loop and 
identify these as \textit{input}s:
every \texttt{source[i]} read inside the loop;
the initial value of \texttt{i} defined outside the loop;
\texttt{max} defined outside the loop; 
and \texttt{first} defined outside the loop.
For cross-iteration redundancy analysis, the \textit{input}s for each
iteration is slightly different:
value \texttt{i} defined in previous iteration, 
the \texttt{source[i]} read in that iteration, 
and two values defined outside the loop, \texttt{max} and \texttt{first}.

Take the big loop at line 7 of Figure~\ref{fig:Apache34464} as an example.
The \textit{input}s for the whole loop include:
three values defined outside the loop
(\texttt{max}, \texttt{first}, \texttt{targetLen}) and every
\texttt{source[.]} and \texttt{target[.]} read inside the loop.
This set is already smaller than the set of all data read by the loop.
We will further shrink this set, removing invariant and others, 
in Section \ref{sec:perf}.
}

@@

After identifying \textit{inputs} using static analysis, 
\Tool instruments the program to record \textit{inputs} values at run time.
%Specifically, we will assign
%a unique ID for each source instruction, and a pair of $<InstID, Value>$
%will be recorded at run time with the execution of a source instruction.
The run-time trace also includes information that
differentiate values recorded from different instructions, loop
iterations, loop instances, etc.
%TODO
%All values defined outside the loop is recorded in block O. 
%We record a delimiter, loop instance number and values defined in previous iterations in block H1'.
%Since we only consider redundant iterations from the same loop instance,  
%we do not record values defined outside the loop for cross-iteration redundancy. 
Once the trace under problematic workload is collected either during
off-line debugging or production runs, 
\Tool will process the trace and decide
whether the loops under study contain cross-iteration or 
cross-loop redundancy. 
%We will first present our high-level algorithms, followed
%by the exact implementation.

\paragraph{High-level algorithms}
We need to answer two questions:
how to judge whether
two iterations or loop-instances are doing redundant work, and
how to judge whether a loop contains sufficient redundant
computation to be inefficient.
Our answer stick to one principle: there should be
sufficient amount of redundant computation to make a loop likely culprit
for a user-perceived performance problem and to make itself worthwhile to
get optimized by developers.

For the first question, \Tool takes a strict 
definition for cross-iteration redundancy and a looser definition for
cross-loop redundancy checking. Specifically,
two iterations that conduct exactly the same computation are
considered redundant; two loop instances that conduct largely
the same computation are considered redundant.
The rationale is that 
a whole loop instance contains a lot of computation, much more than
one iteration in general. Even if only part of its computation
is redundant, it could still be the root-cause of a user-perceived performance
problem and worth developers' attention. 
In practice, we rarely see different loop 
instances doing exactly the same computation.
In cross-loop redundancy examples shown in 
Figure~\ref{fig:Mozilla477564} and
Figure~\ref{fig:Apache34464}, each loop instance is doing similar, but not
exactly the same, work from its previous instance.
For the second question, we believe there should be a threshold. 
A loop will be considered inefficient if its \textit{redundancy rate} goes
beyond the threshold.

\comment{
In our current 
prototype,
when the number of distinct iterations is less than half of the total iterations, 
we consider the loop to be cross-iteration redundant. 
}
%when the number of distinct iterations is less than half of the total iterations 
%iterations/(distinct iterations) = 2

@@

For example, Figure~\ref{fig:Apache37184} shows a loop from Apache-Ant that
contains cross-iteration redundancy.
Under the problem triggering input, there are
only few distinct listeners, and most of the
loop iterations in \texttt{fireMessageLoggedEvent}
are doing redundant work. 
}
%The problem for this bug is that developers say that the performance 
%loss happens in fireMessageLoggedEvent, but in the code I found, function
%messageLogged does not contain anything, and it is an empty function.
%I guess either developers do not have correct understanding of this bug,
%or I did not find the correct codes.

%should they contain exactly the same number of iterations and doing exactly
%the same computation in each iteration?
%Second, is a (static) loop problematic when it contains two instances that
%are redundant with each other?

@@

The key of this implementation is
to judge whether two dynamic loop instances $l_1$ and $l_2$ 
are redundant or not.
The challenge is that $l_1$ and $l_2$ may have executed different numbers of
iterations; in different iterations, different sets of input instructions
may have executed. Therefore, we cannot simply chain values from different
input instructions and iterations together and compare two data sequences.
Instead, we decide to check the redundancy for each input instruction across
$l_1$ and $l_2$ first, and then use the average \textit{redundancy rate} of
all input instructions as the \textit{redundancy rate} between
$l_1$ and $l_2$.

We calculate the redundancy for one input instruction $I$ by normalizing the
edit-distance between the two sequences of values returned by $I$ in the two
loop instances, denoted by $SeqA$ and $SeqB$. 
The exact formula is the following:

@@

\vspace{-0.1in}
\[
\Scale[0.8]{
	Redundancy(I) =  \frac{EditDistance(SeqA, SeqB) - |Length(SeqA) - Length(SeqB)|}{max(Length(SeqA), Length(SeqB))}
}
\]

\comment{
Here, $SeqA$ and $SeqB$ represent the two value sequences corresponding to $I$
from two loop instances, with $SeqA$ being the longer sequence.
$dist$ means edit distance, and $len$ means the length of a value sequence.
Since the edit distance is at least the length-difference between the
two sequences and at most the length of the longer sequence, we use the 
subtraction and division shown in the formula above to normalize the
redundancy value.
}

@@

For cross-iteration redundancy analysis,
we randomly decide at the
beginning of every iteration whether to track the values returned by
input instructions in this iteration.
The implementation is similar with previous sampling work 
\cite{liblit03,liblit05}.
Specifically, we create a clone of the original
loop iteration code, 
%including functions called by the loop directly or indirectly, 
and insert value-recording instructions along the
cloned copy. 
We then insert a code snippet that randomly decides to execute the
cloned copy or the original copy at the beginning of a loop iteration. 
\comment{
Two variables \texttt{CurrentID}, which is initialized as 0, 
and \texttt{NextSampleID}, which is initialized by a random integer, 
are maintained
in this code snippet. \texttt{CurrentID} is increased by 1
for each iteration. When it matches \texttt{NextSampleID}, the control
flow jumps to the value-recording clone of the loop iteration and the 
\texttt{NextSampleID} is increased by a random value. Different sampling
sparsity setting will determine the range from which the random value is
generated.
}

@@

For cross-loop redundancy analysis,
We randomly decide at the beginning
of every loop instance whether to track values for this instance. 
Since we will need to compare two consecutive loop
instances for redundancy, once we decide to sample one loop instance, we will
make sure to sample the immediate next loop instance too.
The implementation is straightforward by cloning the whole loop and making
sampling decisions in the loop pre-headers.
%; (2)
%sampling is conducted when \texttt{CurrentID} equals either
%\texttt{NextSampleID} or \texttt{NextSampleID}$+1$, with \texttt{NextSampleID}
%increased by a random value in the latter case.

@@

First, we avoid tracking multiple reads that we can statically 
prove to return the same value.
Since we implement \Tool in LLVM, we leverage the SSA construction and the
\textit{mem2reg} pass in LLVM to avoid unnecessary tracking of
stack variables.
For example, the read of \texttt{max} on line 10 of 
Figure \ref{fig:Apache34464} is an \textit{input} instruction of the 
\texttt{while} loop on the same line. LLVM identifies it as a loop invariant
and lifts the read of \texttt{max}
out of the loop. Consequently, \Tool only records
its value once during each loop, instead of each iteration.
LLVM is conservative in lifting heap/global variables out of
a loop.
%, for fear of changing the location of potential exceptions.
\Tool conducts a best-effort
loop invariant analysis for heap/global variables.
For example, \Tool identifies that the values of
\texttt{aNode->localName} and \texttt{aNode->namespaceURI} do not change
throughout one loop instance
in Figure~\ref{fig:Mozilla477564}, and hence only traces them once
outside the loop. 
\comment{
This analysis is sound but not complete.
\Tool may miss some loop invariants. For in-house debugging, since we know
the problem-triggering inputs do not lead to exceptions, this optimization
is safe. For production-run usage, users can turn off this heap/global
variable related optimization.
}

Second, we leverage the scalar evolution (SE) analysis in LLVM to remove the 
monitoring to some loop-induction related variables.
SE analysis can tell which variables are
loop-induction variables (e.g., \texttt{i} on line 10
of Figure~\ref{fig:Apache34464}) and what are their strides. 
In cross-iteration redundancy checking,
if a loop iteration's input set contains a loop-induction
variable, we know different iterations' inputs would be different and 
hence conclude that there is no redundancy without
run-time analysis.
In cross-loop redundancy checking,
when the address of a read is a loop induction variable, such as
\texttt{source[i]} on line 10 of Figure~\ref{fig:Apache34464},
we only record the address range at run time,
instead of every value returned by the memory read.
This optimization could lead to false positives ---
different loop instances may work on variables read from
similar memory locations but with different values.
We did not encounter such false positives in our experiments.

\subsection{Fix Strategy Recommendation}
\label{sec:redundant_fix}

\Tool suggests a fix strategy for each inefficient loop 
and provides related information for carrying out that strategy.
This process is straightforward following our taxonomy discussion in 
Section \ref{sec:study}, as there is often only one or two natural fix 
strategies for each sub-category of root causes.

Once a loop is identified as \emph{resultless}, \Tool will suggest
the loop to be deleted in case of 0$^*$,
conditionally skipped in case of [0$|$1]*, or a
data-structure change in case of 0*1?.
\Tool also reports side-effect instructions 
and what percentage of loop iterations executes each such instruction,
helping developers refactor the loop or change data structures.

When a loop is identified as \emph{redundant}, \Tool conducts
extra analysis to
decide whether batching or memoization should be used. 
For cross-iteration redundancy, \Tool suggests batching if the
redundancy is related to I/O operations and suggests memoization otherwise.
Specifically, when the only side effect of a loop is from I/O operations
and the same statement(s) is executed in every iteration, \Tool suggests
batching the specific I/O operations.
Otherwise, \Tool suggests memoization, and reports
input instructions that lead to redundant computation and redundancy rate.

For cross-loop redundancy, whether to use memoization or batching often
depends on which strategy is cheaper.
\Tool uses a simple
heuristic. If 
the side effect of each loop instance is to update 
a constant number of memory locations, like the 
buggy loop in Figure~\ref{fig:Mozilla477564} and Figure~\ref{fig:Apache34464}, 
we recommend memoization. If the side effect is to update
a sequence of memory locations, with the number of locations increasing
with the workload, memoization is unlikely to save much and
hence batching is suggested.
\Tool will also report
the memory reads that return the same values
again and again, which can help decide what to memoize.

### 2016-08-25 section/evaltax.tex
shrink through section 3 and 4

\input section/tab_appbug
\input section/tab_root

The root-cause taxonomy presented in Section \ref{sec:study} is the foundation
of \Tool. In this section,
we quantitatively 
assess the coverage, actionability, and generality of our taxonomy using
a set of
real-world inefficient loop problems collected by previous work
\cite{SongOOPSLA2014,PerfBug}.

@@

Previous work \cite{PerfBug,SongOOPSLA2014} studied the on-line bug
databases of five representative open-source software projects, as 
shown in Table \ref{tab:app_bug}. Through a mix of random sampling and 
manual inspection, they 
found 65 performance problems that are perceived and reported by users. 
Among these 65 problems, 45 problems are related to inefficient loops and 
hence are the target of the study 
here\footnote{The definition of ``loop-related'' in this paper is a little
bit broader than earlier paper~\cite{SongOOPSLA2014}, which only considers
43 problems as loop-related. }.
More details can be found in previous papers that collected
these bugs.

@@

\paragraph{Coverage}
As shown in Table \ref{tab:root}, 
our taxonomy does cover all inefficient loops under study. 
Resultless loops are about as common as redundant loops
(24 vs. 21).
Not surprisingly, 0* loops
are rare in mature software. In fact, no bugs in this
benchmark suite belong to this category.
All other root-cause sub-categories are well represented.

\paragraph{Actionability}
As demonstrated in Table \ref{tab:root}, 
the root-cause categories in our taxonomy are well correlated with
fix strategies.
This indicates that our taxonomy is actionable --- once the root cause
is identified, developers roughly know how to fix the problem.
For example, 
almost all 0*1? resultless loops are fixed by data-structure changes;
all [0$|$1]* resultless loops are
fixed by conditionally skipping the loop;
almost all redundant loops are fixed either by 
memoization or batching. 
The only problem is that there are no silver bullets for fixing 1* loops.
%all inefficient loops with xx root cause are fixed by xxx.
%xxx

\paragraph{Generality}
The root-cause categories in our taxonomy are designed to be 
generic. Table \ref{tab:root} also shows that these categories 
each appears in multiple
application in our study. The only exception is $0^*$-resultless, which
never appears.

@@

\subsection{Caveats} 
What presented above reflects our best effort in assessing our root-cause
taxonomy using a non-biased set of real-world inefficient loop problems that
have been perceived by users and fixed by developers.

@@

Previous work~\cite{PerfBug, SongOOPSLA2014} uses developers tagging and
on-line discussion to judge whether a bug report is about
performance problems and whether 
the performance problem under discussion is noticed and reported by users
or not.
We follow the methodology used in previous work~\cite{SongOOPSLA2014} to judge 
whether the root-cause of a performance problem is related to loops or not.
We do not intentionally ignore any aspect of loop-related performance problems. 
Some loop-related performance problems may never be noticed by end users
or never be fixed by developers, and hence skip our study. However,
there are no conceivable ways to study them.

>>>>

%\input section/tab_appbug
\input section/tab_root

@@

Previous work \cite{PerfBug,SongOOPSLA2014} studied the on-line bug
databases of five representative open-source software projects:
(1) Apache suite, including HTTPD web server written in C/C++,
TomCat server written in Java, and Ant build management tool in Java;
(2) Chromium Google Chrome browser in C/C++;
(3) GCC compiler suite in C/C++;
(4) Mozilla suite, including Firefox web browser and Thunderbird
Email client in C/C++ and JavaScript; 
(5) MySQL database server in C/C++.
Through a mix of random sampling and 
manual inspection, they 
found 65 performance problems that are perceived and reported by users. 
Among these 65 problems, 45 problems are related to inefficient loops and 
hence are the target of the study 
here.
\comment{
\footnote{The definition of ``loop-related'' in this paper is a little
bit broader than earlier paper~\cite{SongOOPSLA2014}, which only considers
43 problems as loop-related. }.
More details can be found in previous papers that collected
these bugs. 
}

@@

\textit{Coverage:}
As shown in Table \ref{tab:root}, 
our taxonomy does cover all inefficient loops under study. 
Resultless loops are about as common as redundant loops
(24 vs. 21).
Not surprisingly, 0* loops
are rare in mature software. 
%In fact, no bugs in this
%benchmark suite belong to this category.
All other root-cause sub-categories are well represented.

\textit{Actionability:}
As shown in Table \ref{tab:root}, 
the root-cause categories in our taxonomy are well correlated with
fix strategies.
This indicates that our taxonomy is actionable --- once the root cause
is identified, developers roughly know how to fix.
For example, 
almost all 0*1? resultless loops are fixed by data-structure changes;
all [0$|$1]* resultless loops are
fixed by conditionally skipping the loop;
almost all redundant loops are fixed either by 
memoization or batching. 
The only problem is that there are no silver bullets for fixing 1* loops.
%all inefficient loops with xx root cause are fixed by xxx.
%xxx

\textit{Generality:}
The root-cause categories in our taxonomy are designed to be 
generic. Table \ref{tab:root} also shows that these categories 
each appears in multiple
application in our study. The only exception is $0^*$-resultless, which
never appears.

@@

\subsection{Threats to Validity} 
What presented above reflects our best effort in assessing our root-cause
taxonomy using a non-biased set of real-world inefficient loop problems that
have been perceived by users and fixed by developers.

@@

Previous work~\cite{PerfBug, SongOOPSLA2014} uses developers tagging and
on-line discussion to judge whether a bug report is about
performance problems and whether 
the performance problem under discussion is noticed and reported by users
or not.
%We follow the methodology used in previous work~\cite{SongOOPSLA2014} to judge 
%whether the root-cause of a performance problem is related to loops or not.
We do not intentionally ignore any aspect of loop-related performance problems. 
Some loop-related performance problems may never be noticed by end users
or never be fixed by developers, and hence skip our study. However,
there are no conceivable ways to study them.
